Network Concurrency Control Done In Band

### 2015-05-30 inband-sync.tex
background

%In this section, we recall the basic SDN principles, with a focus on
%OpenFlow~\cite{of-spec}, the standard SDN protocol today.

\noindent\textbf{Control and data planes}
Software-defined networking decouples the \emph{control plane}, software
controllers that configure the network, from the \emph{data plane}, the
devices that implement the network.
%
In OpenFlow, the \emph{de facto} standard SDN protocol
today, the data plane consists of \emph{OpenFlow switches}.
In a nutshell, the configuration of an OpenFlow switch is a set of
\emph{rules}.
A rule is essentially a \emph{match-action} pair:
the match part of a rule identifies the space of packet headers that are
subject to the rule (e.g., all packets to a specific destination) and
the action part identifies how the switch should process the matched
packets (e.g., forward them to a specific port).

More precisely, the match part of a rule is
a ternary pattern over packet \emph{header} fields.
The pattern is represented as a \emph{value} and a \emph{mask}. \liron{we need to discuss about this, I am not sure this description is accurate.}
In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
addresses, can be \emph{wildcarded}, stipulating that their content does
not affect the match.
Certain fields, such as \emph{ipv4\_src }, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
bitmasked.
In our implementations, we are going to use bitmasking of the metadata
field for performing conditional operations on  its content.
For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.

A \emph{configuration} of an OpenFlow switch is represented as a
collection \emph{flow table}.
A flow table is a set of \emph{flow entries}, each containing a rule,
with a match, an action, and a \emph{priority} level, and flow entries
are ordered according to their priorities.
In the simplest setting, a switch maintains one flow table (table 0).
A data plane packet arriving at a switch is first checked against
the rule with the highest priority and in table $0$.
If the header of the packet fits the match fields of that rule,
a default \emph{instruction} associates the packet with the corresponding action.
Otherwise, the packet is checked against flow  entries with lower
priorities, and if no matching rule is found, the packet is dropped.

@@

%\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
%\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}

\noindent\textbf{Interaction between the planes: FlowMod commands}
The flow tables installed on the switches across the data plane
determine the network \emph{policy}.
Controllers can change the policy by sending
control messages containing \emph{FlowMod commands}.
In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
a modification to an existing flow entry.
%
%
%[[PK already said
%A flow entry consists of: (1)~a match
%field
%(a value and a mask), (2)~a set of actions, and (3)~a priority.
%]]
%
%
A FlowMod command can either add a flow entry or delete a flow entry.
The standard processing of a FlowMod $\add$ command received by a switch is
as follows.
If the switch already maintains a flow entry with \emph{exactly} the
same match and priority level, then the new flow entry will simply \emph{replace} it.
Otherwise, the flow entry will be installed in addition to existing
ones.
A $\dele$ command simply removes an existing flow entry with the same
match and priority,  or does nothing if such an entry is not present.
\petr{Is this correct?}

@@

If rules with overlapping but not identical match fields have the same priority,
inconsistencies can arise, since there is no clearly specified action
that should be applied to a packet matching the two entries.
In order to avoid such conflicts, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
if the switch maintains a flow entry with an overlapping but not
identical match part with the same priority but a different action,
 and if the two rules have the same priority, then
\emph{FlowMod} will fail.
The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
conflicts between controllers, and will be instrumental
in implementing our synchronization abstractions.

%[[PK not sure this is the right place
%implement synchronization objects which form
%the basis of our transactional interface.
%]]
We will use the following notation to create a \emph{FlowMod} message:
$\FlowMod(\match, op, \action, \flags)$, where
%$\match$ denotes the match,
$op$ is $\add$ or $\dele$, and $\flags$ are used, e.g.,  to activate the $\checko$
feature.
A standard \emph{FlowMod} message (and also the flow entries) may specify
more parameters, which are however irrelevant for our implementation, and we suggest to
use the default values.

@@

\noindent\textbf{Bundling}

Another important feature in OpenFlow is \emph{bundling} (as of OpenFlow version 1.4). \stefan{liron, is that correct?}
A controller can send multiple $\FlowMod$ commands  equipped with
the same \emph{bundle identifier}.
These commands are buffered temporarily by
the switch until a \emph{bundle commit} message is received from the
controller.
The bundled commands are then performed with \emph{all-or-nothing semantics}:
either all of them are performed, or none of them is.
In particular, if a configuration request contained in one of the bundled commands cannot
be applied (e.g., the command is rejected because of the set
$\checko$ flag and a conflicting flow entry), all  other commands in
the bundle are rejected and an error
message is sent to the controller that issued them.
%If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
The error message contains $\xid$, the identifier of the first failed command in the
bundle, and $\ecode$, the \emph{error code} corresponding to the failure.

>>>>

%In this section, we recall the basic SDN principles, with a focus on
%OpenFlow~\cite{of-spec}, the standard SDN protocol today.
\hspace{1mm}
\noindent\textbf{Control and data planes.} Software-defined networking
decouples the \emph{control plane}, software
controllers that configure the network, from the \emph{data plane}, the
devices that implement the network.
%
In OpenFlow, the \emph{de facto} standard SDN protocol
today, the data plane consists of \emph{OpenFlow switches}.
In a nutshell, the configuration of an OpenFlow switch is a set of
\emph{rules}.
A rule is essentially a \emph{match-action} pair:
the match part of a rule identifies the space of packet headers that are
subject to the rule (e.g., all packets to a specific destination) and
the action part identifies how the switch should process the matched
packets (e.g., forward them to a specific port).

More precisely, the match part of a rule is
a ternary pattern over packet \emph{header} fields.
The pattern is represented as a \emph{value} and a \emph{mask}. 
%\liron{we need to discuss about this, I am not sure this description is accurate.}
In the mask, certain header fields, e.g., TCP or UDP port numbers or destination and source
addresses, can be \emph{wildcarded}, stipulating that their content does
not affect the match.
Certain fields, such as \emph{ipv4\_src}, \emph{ipv4\_dst}, \emph{metadata}, can be arbitrarily
bitmasked.
In our implementations, we are going to use bitmasking of the \emph{metadata}
field for performing conditional operations on  its content.
%For example, \texttt{ipv4\_src=10.1.14.0, ff.ff.ff.f0} matches all packets
%with source IP addresses in the range \texttt{10.1.14.0}--\texttt{10.1.14.127}.

A \emph{configuration} of an OpenFlow switch is represented as a
a collection of \emph{flow tables}.
A flow table is a set of \emph{flow entries}, each containing a rule,
with a match, an action, and a \emph{priority} level, and flow entries
are ordered according to their priorities.
In the simplest setting, a switch maintains one flow table (table 0).
A data plane packet arriving at a switch is first checked against
the rule with the highest priority and in table $0$.
If the header of the packet fits the match fields of that rule,
a default \emph{instruction} associates the packet with the corresponding action.
Otherwise, the packet is checked against flow  entries with lower
priorities, and if no matching rule is found, the packet is dropped.

@@

%\liron{I think that maybe the rest of this section should be explained by the following subsections, providing in depth description of the OpenFlow features.}
%\liron{Another direction is to have a simplified explenation as now and to refference spesific sections of the standard in side our paper.}
\hspace{1mm}
\noindent\textbf{Interaction between the planes: FlowMod commands}
The flow tables installed on the switches across the data plane
determine the network \emph{policy}.
Controllers can change the policy by sending
control messages containing \emph{FlowMod commands}.
In a nutshell, a \emph{FlowMod} command either specifies a new flow entry or
a modification to an existing flow entry.
%
%
%[[PK already said
%A flow entry consists of: (1)~a match
%field
%(a value and a mask), (2)~a set of actions, and (3)~a priority.
%]]
%
%
A FlowMod command can either add a flow entry or delete a flow entry.
The standard processing of a FlowMod $\add$ command received by a switch is
as follows.
If the switch already maintains a flow entry with \emph{exactly} the
same match and priority level, then the new flow entry will simply \emph{replace} it.
Otherwise, the flow entry will be installed in addition to existing
ones.
A $\dele$ command simply removes an existing flow entry with the same
match and priority,  or does nothing if such an entry is not present.
%\petr{Is this correct?}

@@

%If rules with overlapping but not identical match fields have the same priority,
%inconsistencies can arise, since there is no clearly specified action
%that should be applied to a packet matching the two entries.
In order to avoid inconsistencies caused by different rules with
overlapping match fields, a \emph{FlowMod} command can be equipped with a the $\checko$ flag:
if the switch maintains a flow entry with an overlapping but not
identical match part with the same priority but a different action,
 and if the two rules have the same priority, then
\emph{FlowMod} will fail.
The $\checko$ flag in a \emph{FlowMod} command helps resolving simple
conflicts between controllers, and will be instrumental
in implementing our synchronization abstractions.

%[[PK not sure this is the right place
%implement synchronization objects which form
%the basis of our transactional interface.
%]]
We will use the following notation to create a \emph{FlowMod} message:
$\FlowMod(\match, op, \action, \flags)$, where
%$\match$ denotes the match,
$op$ is $\add$ or $\dele$, and $\flags$ are used, e.g.,  to activate the $\checko$
feature.
For simplicity, we omit other standard parameters, assuming them to carry
default values.
%A standard \emph{FlowMod} message (and also the flow entries) may specify
%more parameters, which are however irrelevant for our implementation, and we suggest to
%use the default values.

@@

\hspace
\noindent\textbf{Bundling.}
%
Another important feature in OpenFlow is \emph{bundling}. %(as of OpenFlow version 1.4). \stefan{liron, is that correct?}
A controller can send multiple $\FlowMod$ commands  equipped with
the same \emph{bundle identifier}.
These commands are buffered temporarily by
the switch until a \emph{bundle commit} message is received from the
controller.
The bundled commands are then performed with \emph{all-or-nothing semantics}:
either all of them are performed, or none of them is.
In particular, if a configuration request contained in one of the bundled commands cannot
be applied (e.g., the command is rejected because of the set
$\checko$ flag and a conflicting flow entry), all  other commands in
the bundle are rejected and an error
message is sent to the controller that issued them.
%If multiple \emph{FlowMod} messages with the $\checko$ flag are sent as a single bundle,
The error message contains $\xid$, the identifier of the first failed command in the
bundle, and $\ecode$, the \emph{error code} corresponding to the failure.
